\chapter{Limitations, Reproducibility, and Future Work}

\section{Limitations}

The COMSOL model is a physics-based accelerated degradation study, not a final
first-principles lifetime prediction for a specific commercial device. The
baseline electronic state must still be calibrated against the measured
fresh-device J--V range, and the degradation rates are accelerated for
computational practicality. The reported aging time should therefore be
interpreted as equivalent accelerated aging time, not as a direct field-time
prediction.

The second limitation is parameter non-uniqueness. Trap growth, shunt leakage,
contact resistance, surface recombination, and photogeneration loss can all
reduce PCE. Several combinations can produce similar terminal J--V degradation.
The model needs hysteresis, impedance, transient-current, optical, and aging
data to separate mechanisms with confidence.

The third limitation is mobile-ion validation. The normalized ion state
$u_{\mathrm{ion}}$ provides a physically useful way to couple redistribution
into Poisson electrostatics, but the ion mobility, concentration, boundary
conditions, and ion--trap coupling need experimental support before they can be
treated as device-specific material parameters. DIT and capacitance-based
measurements provide stronger constraints than terminal J--V curves, but the
integrated current must be baseline-corrected so that electronic plateaus and
capacitive spikes are not interpreted as mobile ionic charge.

The fourth limitation is implementation consistency with the laboratory stack.
The physical architecture is
glass/ITO/NiO$_x$/MeO-2PACz/perovskite/C$_{60}$/BCP/Ag/UV-resin/glass,
illuminated from the ITO side. Historical COMSOL objects retain FTO/Au labels
and GUI-history assignments that may use a reversed layer coordinate. Those
labels, selections, illumination direction, and contact work functions must be
corrected and re-solved before the model is called device-specific.

The fifth limitation is export completeness. The current canonical
light-temperature stress-matrix J--V export is physically scaled and usable for
PCE and $J_{sc}$ across all 360 curves, but its 1.25 V diagnostic voltage limit
does not resolve zero-current crossings for all mild or early-aging cases.
Quantitative $V_{oc}$ and FF claims therefore remain limited to the subset of
curves that cross zero current until a wider diagnostic voltage sweep is
exported.

\section{Reproducibility}

The COMSOL processing workflow is reproducible from the repository root:
\begin{verbatim}
python .\scripts\run_comsol_paper_pipeline.py
\end{verbatim}
The report can then be rebuilt with:
\begin{verbatim}
latexmk -pdf -interaction=nonstopmode main.tex
\end{verbatim}
The processed tables and figures under \path{data/processed/comsol/} and
\path{outputs/} provide a traceable path from exported COMSOL data to the
reported J--V metrics, lifetime markers, and mechanism-sensitivity plots.
The architecture-stress extensions are regenerated with:
\begin{verbatim}
python .\scripts\process_comsol_architecture_stress_jv.py
python .\scripts\process_comsol_profile_grid.py
python .\scripts\plot_comsol_spatial_profiles.py
python .\scripts\plot_comsol_band_diagrams.py
\end{verbatim}

\section{Data Availability}

The thesis source, processing scripts, generated figures and tables, selected
COMSOL exports, and public PDF are maintained in the supporting public
repository at \url{https://github.com/flabio121/thesis-public}. Large raw
COMSOL intermediates and local model files are excluded from the public copy
when they are too large for practical version control or contain active local
working files. Their generation paths and provenance are documented in
Appendix C, \path{REPRODUCIBILITY.md}, and the traceability matrix.

\section{Future Work}

The most important next step is improved calibration. The pristine COMSOL
baseline should be fit to the same device architecture and measurement
conditions as the experimental J--V data. The fit should match $V_{oc}$,
$J_{sc}$, FF, and PCE before degradation parameters are interpreted.

The second step is implementation and validation of light-soaking and
self-healing terms. The proposed conditioning state can represent improved
contact extraction under illumination, while optional healing terms can
represent partial recovery during rest. These terms require controlled
light-soaking, dark-rest, and repeated-scan experiments before they should be
used in predictive simulations.

The third step is ion-sensitive validation. EIS, TPC, capacitance, ion-current,
and scan-rate-dependent hysteresis measurements can constrain
$c_{\mathrm{ion},0}$, $\mu_{\mathrm{ion}}$, $D_{\mathrm{ion}}$, boundary
conditions, and ion--trap coupling. These measurements are needed because
terminal J--V curves alone do not uniquely determine the mobile-ion state.
Future DIT simulations should pair each mobile-ion case with a frozen-ion
control so that $I_{\mathrm{corr}}(t)$ and the extracted $N_0$ are defined by
the ionic transient rather than by the full terminal current.

The fourth step is architecture-specific parameterization. The model should be
calibrated separately for device stacks, contacts, and barrier layers because
the same degradation pathway can produce different electrical signatures in
different architectures.

The fifth step is completion of the spatial and transient result set. The
selected exports now support band diagrams with quasi-Fermi levels, the
correct HTL-to-ETL plotting direction, and selected ion/trap line-profile
panels from the profile grid. Final validation should still include explicit
short-circuit, maximum-power, and open-circuit band diagrams, a strict
COMSOL-operator ion-conservation check, and frozen-ion-controlled DIT charge
extraction. The quantitative DIT result remains a future validation step until
a frozen-ion or no-ion control current is exported for baseline subtraction.

The final step is a COMSOL-calibrated surrogate workbench. A Streamlit
prototype exists as supporting infrastructure for this thesis workflow, and
the current 6 $\times$ 6 light-temperature export now supports a first
traceable lite surrogate baseline. That baseline demonstrates grouped
COMSOL-to-COMSOL interpolation for PCE, normalized PCE, PCE retention, $J_{sc}$,
and J--V current density, but it is still forward-only and calibrated only to
the present COMSOL grid. The current lite surrogate results are generated by
\path{scripts/train_lite_surrogates.py}; the Streamlit app remains a prototype
interface rather than the production deployment of those models. The web app
still needs uncertainty estimates, active-learning run selection, inverse
diagnosis from measured J--V history, and matched experimental calibration. In
the intended final role, the surrogate would accelerate parameter exploration,
identify
high-value next COMSOL simulations, and support latent-state inference from
measured J--V history. It would not replace the COMSOL drift--diffusion model
as the physics source of truth.

\section{Concluding Perspective}

This thesis develops a COMSOL-based degradation modeling framework for
perovskite solar cells. The model is a physics-based platform for studying
coupled electronic, ionic, and degradation behavior. The implementation
demonstrates qualitative mechanism separation through drift-diffusion transport,
mobile-ion electrostatics, trap-density growth, contact resistance, shunt
leakage, voltage protocols, Beer--Lambert optical generation, DIT pulse
analysis, electronic parameter sweeps, and accelerated J--V aging studies.
Experimental calibration remains
ongoing, and future work should focus on improved calibration, self-healing and
light-soaking implementation, ion-sensitive validation, architecture-specific
parameters, and broader COMSOL-calibrated surrogate use after the data set is
physically grounded.
